### 2024-11-14 paper/paper.qmd
Initial setup

---
title: "Analysing the 2024 U.S. Presidential Election"
subtitle: "Spotlight on North Carolina's Role as a Battleground State"
author: 
  - David Qi
  - Haobo Ren
thanks: "Code and data are available at: https://"
date: today
date-format: long
abstract: ""
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(knitr)
library(ggplot2)
```

# Introduction {#sec-intro}

# Data {#sec-data}

```{r}
#| label: tbl-polloverview
#| tbl-cap: Comparison of average polling result for two candiates
#| include: true
#| echo: false
#| warning: false
#| message: false

merged_summary = merged_summary %>% select(state,total_count.x,mean_pct_harris,mean_pct_trump,abs_diff)  %>% arrange(abs_diff)

kable(merged_summary, col.names = c("State", "Number of Polls" , "Average Support Rate(Harris)","Average Support Rate(Trump)","Absolute difference"))
```

# Results {#sec-results}

# Discussion{#sec-discussion}

\newpage

\appendix

\newpage

# References