### 2024-11-15 paper/paper.qmd

---
title: "Analysing the 2024 U.S. Presidential Election"
subtitle: "Spotlight on North Carolina's Role as a Battleground State"
author: 
  - David Qi
  - Haobo Ren
thanks: "Code and data are available at: https://"
date: today
date-format: long
abstract: ""
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(knitr)
library(ggplot2)
```

# Introduction {#sec-intro}

# Data {#sec-data}

## Mesurement
The data used in this paper were obtained from the websites of various food retail companies, including Voila, T&T, Loblaws, No Frills, Metro, Galleria, Walmart Canada, Save-On-Foods. The data was gathered by using screen-scrape of website UI, starting from Feb.28 to Nov.14. This data captures the product name, units, and prices of each vendors.

This data has some limitations. There maybe missing data on certain days for certain vendors, days on which those specific extracts failed. This data didn't include all vendors in the markets, it may cause bias to our analysis.

```{r}
#| label: tbl-polloverview
#| tbl-cap: Comparison of average polling result for two candiates
#| include: true
#| echo: false
#| warning: false
#| message: false

merged_summary = merged_summary %>% select(state,total_count.x,mean_pct_harris,mean_pct_trump,abs_diff)  %>% arrange(abs_diff)

kable(merged_summary, col.names = c("State", "Number of Polls" , "Average Support Rate(Harris)","Average Support Rate(Trump)","Absolute difference"))
```

# Results {#sec-results}

# Discussion{#sec-discussion}

\newpage

\appendix

\newpage

# References