### 2024-11-15 paper/paper.qmd
Added Figure 1 and 2, updated data section

---
title: "Analysing the 2024 U.S. Presidential Election"
subtitle: "Spotlight on North Carolina's Role as a Battleground State"
author: 
  - David Qi
  - Haobo Ren
  - Xinrui Xie
  - Ruikang Wang
  - Sakura Hu
thanks: "Code and data are available at: https://github.com/UTDQi/grocery_analysis"
date: today
date-format: long
abstract: ""
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(knitr)
library(ggplot2)
library(janitor)

tea_summary = read_csv(here::here("data/02-analysis_data/tea_summary.csv"))

price_change = read_csv(here::here("data/02-analysis_data/price_change_lemon_ginseng.csv"))
```

# Introduction {#sec-intro}

# Data {#sec-data}

## Mesurement
The data used in this paper were obtained from the websites of various food retail companies, including Voila, T&T, Loblaws, No Frills, Metro, Galleria, Walmart Canada, Save-On-Foods. The data was gathered by using screen-scrape of website UI, starting from Feb.28 to Nov.14. This data captures the product name, units, and prices of each vendors. [@raw_data_source]

This data has some limitations. There are missing data on certain days for certain vendors,some missing points occur on days which those specific extracts failed. While data collection for some products starts form July. This data didn't include all vendors in the markets, which may cause bias to our analysis.

## Key Observation Variables

### Current Time
The current time(`current_price`) is the time when this data is retrieved.

### Current Price
The current price(`current_price`) is the price of the product retrieved at the time accessed.

### Old Price
The current price(`old_price`) is the "old price" of the product retrieved at the time accessed that is stuck out by the vendor. It appears whenever there is a sales on and the vendor wants to show the change in price. This is the key variable to differentiate between sale and silent price drops.

## Overview
[@fig-PriceVsVendor] is a graph

```{r}
#| label: fig-PriceVsVendor
#| fig-cap: Distribution of Average Prices by Vendor
#| include: true
#| echo: false
#| warning: false
#| message: false

tea_summary = clean_names(tea_summary)

ggplot(tea_summary, aes(x = avg_price, fill = vendor)) +
  geom_histogram(binwidth = 1, alpha = 0.7) +
  labs(
       x = "Average Price",
       y = "Count") +
  facet_wrap(~ vendor, scales = "free_y") +
  theme(
    legend.position = "none", 
    plot.margin = margin(t = 15, r = 15, b = 20, l = 15)  # Increase bottom margin
  )
```

```{r}
#| label: fig-ChangeInGinsengLemon
#| fig-cap: Price Change of Lipton Ginseng Lemon Green Tea 72 Counts
#| include: true
#| echo: false
#| warning: false
#| message: false

price_change = clean_names(price_change)

# Ensure `nowtime` is in datetime format and `current_price` is numeric
price_change$nowtime <- as.POSIXct(price_change$nowtime)
price_change$current_price <- as.numeric(price_change$current_price)

# Sample ggplot code to create the point graph
ggplot(price_change, aes(x = nowtime, y = current_price, color = vendor)) +
  geom_point(alpha = 0.6) +
  labs(x = "Time(Month)",
       y = "Current Price",
       color = "Vendor") +
  theme_minimal() 
```

# Results {#sec-results}

# Discussion{#sec-discussion}

## Correlation Versus Causation

## Missing Data

## Sources of Bias.

\newpage

\appendix

\newpage

# References